---
title: "Ch 10"
output: html_document
---

Question(7)
```{r}
library(ISLR)
set.seed(1)
uscale=scale(USArrests)
corr_dist = as.dist(1-cor(t(uscale)))
eucl_dist = dist(uscale)^2
summary(corr_dist/eucl_dist)
```

Question(8)
part(a)
```{r}
set.seed(1)
pr.out = prcomp(USArrests, scale=TRUE)
prvar = pr.out$sdev^2
pve_1 = prvar/sum(prvar) 
pve_1
```


part(b)
```{r}
scaled_data = scale(USArrests)
head(pr.out$x==(scaled_data %*% pr.out$rotation)) # To check if calculations correct
```

```{r}
den = sum(apply(scaled_data^2, 2, sum)) # denominator for eq. 10.8

```



```{r}
# Function to calculate numerator of eq 10.8
var_exp = function(data, rotation){
  dim1 = dim(data)[1]
  dim2 = dim(data)[2]
  vars = rep(0, dim2)
  for (m in 1:dim2){
    sum = 0
    for (i in 1:dim1){
      sum = sum + (data[i,] %*% rotation[,m])^2
      
    }
    vars[m] = sum
  }
  return(vars)
}
  
num = var_exp(scaled_data, pr.out$rotation)
pve_2 = (num / den)
print(pve_2)
print(pve_1)
isTRUE(all.equal(pve_2, pve_1))
```

Values are same using two methods.



Question(9)
part(a)
```{r}
hc.complete=hclust(dist(USArrests))
plot(hc.complete, main="Complete Linkage", xlab="", ylab="", cex=0.8)
```

part(b)
```{r}
ct = cutree(hc.complete, k=3)
for (i in 1:3){
  print(i)
  print(rownames( USArrests )[ct == i])
  cat("\n\n")
}
```


part(c)
```{r}
scaled_data = scale(USArrests)
hc.compsc=hclust(dist(scaled_data))
plot(hc.compsc, main="Complete Linkage (Scaled Version)", xlab="", ylab="", cex=0.8)
```


```{r}
ct = cutree(hc.compsc, 3)
for (i in 1:3){
  print(i)
  print(rownames( USArrests )[ct == i])
  cat("\n\n")
}
```


```{r}
table(cutree(hc.compsc, 3), cutree(hc.complete, 3))
```

part(d)

Clusters obtained from cutting tree in 3 have changed significantly as most states belong to third group now.

Yes, variables should be scaled as the UrbanPop is % of population but other variables measure cases out of 1,00,000 population which is quite different value.




Question(10)
part(a)
```{r}
k = 3
n = 20
p = 50
set.seed(1)
dataset = rnorm(c(k*n*p))
dim(dataset)=c(60,50)
dataset[1:20,] = dataset[1:20,] + 0.65
dataset[41:60,] = dataset[41:60,] - 0.65

```


part(b)
```{r}
pr.out=prcomp(dataset, scale=TRUE)
plot(pr.out$x[,1], pr.out$x[,2], col=labels, pch=19)
```


part(c)
```{r}
km.out = kmeans(dataset, centers=3, nstart=25)
km.out
```

```{r}
plot(dataset, col=km.out$cluster, pch=19)
```

```{r}
table(c(rep(3, 20), rep(1,20), rep(2,20)), km.out$cluster) #labelling the way clusters are predicted
```


error rate = 2 / 60 = 3.33%

Clusters compare to true labels quite well

part(d)
```{r}
km.out=kmeans(dataset, centers=2, nstart=25)
km.out
```

```{r}
plot(dataset, col=km.out$cluster, pch=19)
```

One of the class has been split between the two classes almost equallly.


part(e)
```{r}
km.out=kmeans(dataset, centers=4, nstart=25)
km.out
```


```{r}
plot(dataset, col=km.out$cluster, pch=19)
```

Clusters have made 4 groups now.

part(f)
```{r}
km.out=kmeans(data.frame(pr.out$x[,1], pr.out$x[,2]), centers=3, nstart=25)
km.out
```


```{r}
labels = c(rep(1, 20), rep(2,20), rep(3,20))
table(labels, km.out$cluster)
```


Error rate = 2 / 60 = 3.33%



```{r}
plot(pr.out$x[,1], pr.out$x[,2], col=labels, pch=19)
```

part(g)
```{r}
scaled = scale(dataset)
km.out = kmeans(scaled, centers=3, nstart=25)
km.out
```




```{r}
plot(pr.out$x[,1], pr.out$x[,2], col=km.out$cluster, pch=19)
```

```{r}
labels = c(rep(2, 20), rep(3,20), rep(1,20))
table(labels, km.out$cluster)
```

2 observations misclassified, so scaled version performs some worse.


Question(11)
part(a)
```{r}
Genes=read.csv("http://faculty.marshall.usc.edu/gareth-james/ISL/Ch10Ex11.csv", header=F)

```

part(b)
```{r}
dd = as.dist(1-cor((Genes)))
hc.complete=hclust(dd)
plot(hc.complete, main="Complete Linkage with Correlation-Based distance", xlab="", sub="")
```

```{r}
dd = as.dist(1-cor((Genes)))
hc.average=hclust(dd, method="average")
plot(hc.average, main="Average Linkage with Correlation-Based distance", xlab="", sub="")

```


```{r}
dd = as.dist(1-cor((Genes)))
hc.single=hclust(dd, method="single")
plot(hc.single, main="Single Linkage with Correlation-Based distance", xlab="", sub="")
```

Yes, 3 Groups in Average Linkage, 2 Groups in Single and Complete Linkage


```{r}
print( table( predicted=cutree( hc.complete, k=2 ), truth=c( rep(1,20), rep(2,20) ) ) )
```


part(c)
```{r}

predicted=cutree( hc.complete, k=2 ) 

n1 = apply( Genes[ predicted==1, ], 2, length ) # the number of samples (number of patients in each cluster)
n2 = apply( Genes[ predicted==2, ], 2, length )

m1 = apply( Genes[ predicted==1, ], 2, mean ) # the means across the 1000 genes in each cluster
m2 = apply( Genes[ predicted==2, ], 2, mean )

v1 = apply( Genes[ predicted==1, ], 2, var ) # the variances across the 1000 genes in each cluster
v2 = apply( Genes[ predicted==2, ], 2, var )

pooled_variance = sqrt( v1 / n1 + v2 / n2 )

t_value = ( m1 - m2 ) / pooled_variance 
plot( t_value, xlab="gene index", ylab="unpaired t-value" )
```

```{r}
t_value
```




